### 2025-01-20 thesis/main.tex
Add new references to main.bib for pattern recognition and graph neural networks

\documentclass{article}
\usepackage{geometry}
\geometry{
    a4paper,
    left=30mm,
    right=30mm,
    top=30mm,
    bottom=30mm
}

\usepackage{annotate-equations}
\usepackage{graphicx}
\usepackage{wrapfig}
% \usepackage{figure}

\graphicspath{ {./media/} }
\usepackage[square,numbers]{natbib}
\usepackage[nottoc]{tocbibind}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{float}

\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{tikz}
\usepackage{listings}
\usepackage{setspace}
\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{0.98,0.98,0.98}
\lstdefinelanguage{Python}{
	numbers=left,
	numberstyle=\footnotesize,
	numbersep=1em,
	xleftmargin=1em,
	framextopmargin=2em,
	framexbottommargin=2em,
	showspaces=false,
	showtabs=false,
	showstringspaces=false,
	frame=l,
	tabsize=4,
	% Basic
	basicstyle=\ttfamily\small\setstretch{1},
	backgroundcolor=\color{Background},
	% Comments
	commentstyle=\color{Comments}\slshape,
	% Strings
	stringstyle=\color{Strings},
	morecomment=[s][\color{Strings}]{"""}{"""},
	morecomment=[s][\color{Strings}]{'''}{'''},
	% keywords
	morekeywords={import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert},
	keywordstyle={\color{Keywords}\bfseries},
	% additional keywords
	morekeywords={[2]@invariant,pylab,numpy,np,scipy},
	keywordstyle={[2]\color{Decorators}\slshape},
	emph={self},
	emphstyle={\color{self}\slshape},
	%
}

\title{\textbf{Bachelor thesis}\\
Exploration of the influence of graph parameters on the generalization error of Graph Neural Networks}

\author{Author: Wensheng Zhang}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
The goal of this bachelor thesis is to understand empirically how graph parameters/characteristics influence the generalization error of Graph Neural Networks (GNNs).
\end{abstract}

\section{Introduction}\label{sec:intro}
Graph learning became a hot topic in the field of machine learning in recent years. Graph Neural Networks (GNNs), also known as deep learning on graphs, are a class of neural networks that can operate on graph-structured data, such as social networks~\cite{easley2010networks}, biological networks~\cite{barabasi2004network}, and images~\cite{simonovsky2017dynamic}. GNNs use the structure of a graph to perform "message passing", allowing information to propagate across nodes and edges. The key idea is to iteratively aggregate and update node features by exchanging information with local neighboring nodes, which allows GNNs to capture the structural information of the graph. This information is then used to make predictions. GNNs have been demonstrated to be effective in a wide range of tasks, such as node classification~\cite{hamilton2017inductive}, edge prediction~\cite{morselli2021network}, and graph classification~\cite{gilmer2017neural}. Popular GNNs in recent years include Graph Convolutional Networks (GCN)~\cite{kipf2016semi}, Graph Attention Networks (GAT)~\cite{velickovic2020pointer}, and Message Passing Neural Networks (MPNN)~\cite{gilmer2017neural}.

However, the behavior of GNNs is not well understood. Particularly, the relationship between graph parameters/characteristics and the empirical generalization error of GNNs is not clear.  The goal of this bachelor thesis is to understand empirically how graph parameters/characteristics impact the generalization error of GNNs.

Speaking more specifically.
First, I calculate the graph parameters for a set of datasets: Each graph has its own unique properties, such as the average degree. In graph theory, \textit{graph parameter} is used to quantify them. And a dataset consists of multiple graphs, each with its graph parameter. In the case, I calculate the average value of our desired graph parameters over all graphs in the dataset. In this thesis, 57 datasets from the TUDataset~\cite{morris2020tudataset} is used for the calculation. Second, I train different GNN models on each dataset and find that their generalization abilities varied. I show that there is a certain relationship between generalization ability and different graph parameters.

\subsection{Related work}
The following section provides an overview of the related work.

\subsubsection{Graph parameters}

\subsubsection{Generalization error}

\section{Background on learning on graphs}
The following sections provide an overview of the concepts and methods related to graph theory, machine learning, and graph neural networks.

\subsection{Notion}
At first, we need to define some terms with regard to the graph. In graph theory, we assume that $G=(V,E)$ is an undirected graph, where $V$ is the set of nodes and $E  \subseteq \{\{v,w\}\mid v,w \in V , v \neq w\}$ is the set of edges. We use $\mathbb{G}$ to represent the set of all possible graphs. The neighborhood of a node $v$ is denoted as $N(v)$, defined as $N(v) = \{w \mid \{v,w\} \in E\}$. The degree of a node $v$, denoted as $d(v)$, is the number of neighbors of $v$, that is, $d(v) = |N(v)|$. In addition to the nodes and edges that comprise the graph, we can attach a more complex set of data to each node.  A feature vector $x_n\in \mathbb{R}^k$, is attached to a node $v_n$, and in some cases, edge feature vectors are also included. These vectors can be used to characterize graphs or nodes in graphs. The adjacency matrix $A \in \mathbb{R}^{|V| \times |V|}$ is a matrix that represents the connections between nodes in the graph. The element $a_{ij}$ of the adjacency matrix is 1 if there is an edge between node $v_i$ and node $v_j$, and 0 otherwise.

\subsection{Machine learning}
In the following section, we provide an overview of the concepts and methods related to machine learning.

\subsubsection{Supervised learning}
Machine learning can be roughly divided into three categories: supervised learning, unsupervised learning, and reinforcement learning~\cite{bishop2006pattern}.

Supervised learning is a type of machine learning that involves training a model on a labeled dataset. The model is trained to learn the relationship between the input features and the output labels. The goal is to learn a mapping from the input features to the output labels so that the model can make predictions on new, unseen data. One common task in supervised learning is classification, where the goal is to assign each input vector to one of a finite number of classes (categories). In the graph learning context, the predicted output can be a label for each node in the graph, a label for the entire graph, or a label for each edge in the graph, e.g. node classification, graph classification, and link prediction~\cite{velivckovic2023everything}.

\subsubsection{Feed-forward neural networks}

Feed-forward neural networks, also known as multilayer perceptrons (MLPs), are a type of neural network that consists of multiple layers of neurons. The neurons in each layer are connected to the neurons in the next layer, and the output of each neuron is computed as a weighted sum of the inputs followed by a non-linear activation function. The output of the network is computed by passing the input through the network and computing the output of the final layer. The weights of the network are learned using an optimization algorithm such as gradient descent. Feed-forward neural networks are a powerful tool for learning complex patterns in data and have been successfully applied to a wide range of tasks, such as image recognition, speech recognition, and natural language processing.

The function of a feed-forward neural network can be written as

$$ 
    y = f(W^{(L)} \cdot f(W^{(L-1)} \cdot \ldots \cdot f(W^{(1)} \cdot x + b^{(1)}) + b^{(2)}) + \ldots + b^{(L)})
$$

where $x$ is the input vector, $W^{(i)}$ and $b^{(i)}$ are the weights and biases of the $i$-th layer, and $f$ is the activation function. The output $y$ is the predicted output of the network.

\subsubsection{Loss function}
The loss function is a function that measures the difference between the predicted output of a model and the true output. The goal of training a machine learning model is to minimize the loss function, which is done by adjusting the parameters of the model using an optimization algorithm such as gradient descent. The choice of loss function depends on the task at hand. For classification tasks, the cross-entropy loss function is commonly used, while for regression tasks, the mean squared error loss function is often used.

The cross-entropy loss function is defined as

$$
    L(y, \hat{y}) = -\sum_{i} y_i \log(\hat{y}_i)
$$

where $y$ is the true output, $\hat{y}$ is the predicted output, and $i$ is the index of the classes. The cross-entropy loss function is used to measure the difference between the true output and the predicted output of a classification model.

\subsubsection{Activation function}
An activation function is a non-linear function that is applied to the output of a neuron in a neural network. The activation function introduces non-linearity into the network, allowing it to learn complex patterns in the data. There are many activation functions that can be used in neural networks, such as the sigmoid function, the tanh function, and the ReLU function. The ReLU function is one of the most commonly used activation functions in deep learning because it is simple and computationally efficient. The ReLU function is defined as

$$
    \text{ReLU}(\cdot)=\max(0,\cdot)
$$

The ReLU function is used to introduce non-linearity into the network and has been shown to be effective in training deep neural networks.

\subsubsection{Gradient Descent}
Gradient descent is an optimization algorithm used to minimize a function by iteratively moving in the direction of the steepest decrease of the function. The algorithm works by computing the gradient of the function at a given point and then updating the point in the direction of the negative gradient. The update rule is given by

$$
    W^{(t+1)} = W^{(t)} - \alpha \nabla f(W^{(t)})
$$
where $W^{(t)}$ is the parameter vector at iteration $t$, $\alpha$ is the learning rate, and $\nabla f(W^{(t)})$ is the gradient of the function $f$ with respect to the parameters $W^{(t)}$. The learning rate $\alpha$ is a hyperparameter that controls the step size of the update. Gradient descent is a popular optimization algorithm used in machine learning to train neural networks and other models.

\subsubsection{Stochastic Gradient Descent}
Stochastic gradient descent (SGD) is a variant of gradient descent that uses a random sample of the training data to compute the gradient at each iteration. The algorithm works by randomly selecting a mini-batch of data points from the training dataset, computing the gradient of the loss function with respect to the mini-batch, and then updating the parameters based on the gradient. The update rule is given by

$$
    W^{(t+1)} = W^{(t)} - \alpha \nabla f(W^{(t)}, x^{(t)})
$$
where $x^{(t)}$ is a mini-batch of data points from the training dataset. Stochastic gradient descent is more computationally efficient than standard gradient descent and can be used to train large-scale machine learning models.

\subsubsection{Early Stopping}
Early stopping is a technique used to prevent overfitting in machine learning models. The idea is to monitor the performance of the model on a validation dataset during training and stop training when the performance starts to degrade. This is done by keeping track of the validation loss and stopping training when the validation loss stops decreasing or starts to increase. Early stopping is used to prevent the model from memorizing the training data and to obtain a more generalizable model.

\subsubsection{Cross-validation}
Cross-validation is a technique used to evaluate the performance of a machine learning model. The idea is to split the dataset into multiple subsets, or folds, and train the model on one subset while evaluating it on the remaining subsets. This process is repeated multiple times, with each subset used as the test set once. The performance of the model is then averaged over all the folds to obtain an estimate of the model's generalization performance. Cross-validation is used to prevent overfitting and to obtain a more accurate estimate of the model's performance.

\subsubsection{Hyperparameter tuning}
Hyperparameter tuning is the process of selecting the optimal hyperparameters for a machine learning model. Hyperparameters are parameters that are set before the training process begins and control the behavior of the model. Examples of hyperparameters include the learning rate, batch size, number of hidden layers, and number of epochs. Hyperparameter tuning is an important step in the machine learning pipeline and can have a significant impact on the performance of the model. There are many techniques for hyperparameter tuning, such as grid search, random search, and Bayesian optimization. In this thesis, Bayesian optimization is used to tune the hyperparameters of the models.

Bayesian optimization is more efficient than grid search and random search because it uses a probabilistic model to guide the search for the optimal hyperparameters. The idea is to model the objective function (e.g. the validation accuracy of the model) as a Gaussian process and use the model to select the next set of hyperparameters to evaluate. This process is repeated until the optimal hyperparameters are found.

\subsubsection{Optimization algorithm}
An optimization algorithm is an algorithm used to minimize a function by iteratively updating the parameters of the model. The goal is to find the set of parameters that minimizes the loss function. Gradient descent is a common optimization algorithm used in machine learning, but there are many variants of gradient descent that are designed to improve convergence speed and stability. Some popular optimization algorithms include Adam~\cite{kingma2014adam}, RMSprop~\cite{graves2013generating}, and Adagrad~\cite{duchi2011adaptive}. These algorithms use different strategies to update the parameters of the model and can be more effective than standard gradient descent in certain situations.

\subsubsection{Generalization}
A key concept in evaluating model performance or generalization ability in such task is the empirical generalization error $E_{gen}$. This is a measure of how well the model generalizes to unseen data, which can be calculated as the difference between the training accuracy and the test accuracy. It is defined as

$$
    E_{gen} = {Acc_{train}} - {Acc_{test}}
$$
where $Acc_{train}$ is the accuracy of the model on the training dataset, and $Acc_{test}$ is the accuracy of the model on the test dataset. The accuracy $Acc$ on a dataset is defined as the proportion of correctly classified graphs, that is, $Acc = \frac{\text{Number of correctly classified graphs}}{\text{Total number of data points}}$.

The empirical generalization error is also known as the generalization gap~\cite{goodfellow2016deep}. Typically, the learning on the training dataset and test dataset behaves differently. The model may perform well on the training dataset but poorly on the test dataset. By observing the generalization error, we can decide on a stopping criterion for the training process. And the generalization error can be used to determine whether the model is overfitting or underfitting. Furthermore, the generalization error can be influenced by many factors, such as the model architecture, the hyperparameters, and the data properties in the dataset (e.g. Graph parameter).

\subsection{Graph Neural Networks}
Graph Neural Networks (GNNs) are a class of neural networks that can operate on graph-structured data. GNNs use the structure of a graph to perform "message passing", allowing information to propagate across nodes and edges. The key idea is to iteratively aggregate and update node features by exchanging information with local neighboring nodes, which allows GNNs to capture the structural information of the graph. This information is then used to make predictions. GNNs have been demonstrated to be effective in a wide range of tasks, such as node classification, edge prediction, and graph classification. Popular GNNs in recent years include Graph Convolutional Networks (GCN), Graph Attention Networks (GAT), and Message Passing Neural Networks (MPNN).

In the following sections, we provide an overview of the methods related to this thesis.

\subsubsection{Graph Convolutional Networks (GCN)}

Graph Convolutional Networks (GCNs) are a class of neural networks that can operate on graph-structured data. The GCN model $f(X, A)$, as proposed by Kipf and Welling~\cite{kipf2016semi}, is defined as

$$ 
 H^{(l+1)} = \sigma(\tilde{D}^{-\frac{1}{2}}\tilde{A}\tilde{D}^{-\frac{1}{2}}H^{(l)}W^{(l)})
$$

where $H^{(l)}\in \mathbb{R}^{N\times D}$ is the node feature matrix at layer $l$; $H^{(0)} = X$. $\tilde{A} = A + I$ is the adjacency matrix of the graph with self-connections added, $\tilde{D}$ is the degree matrix of $\tilde{A}$, $W^{(l)}$ is the weight matrix of the layer, and $\sigma$ is the activation function, such as the $\text{ReLU}(\cdot)=\max(0,\cdot)$. The GCN layer aggregates information from neighboring nodes and updates the node features by applying a linear transformation followed by a non-linear activation function. The output of the GCN layer is the node feature matrix at the next layer. The GCN layer can be stacked to create deep GCN models, which have been shown to be effective in a wide range of tasks, such as node classification, edge prediction, and graph classification.

\subsubsection{Simplified Graph Convolution (SGC)}

Simplified Graph Convolution (SGC)~\cite{wu2019simplifying} is a simplified version of the GCN model that removes the non-linear activation function and the normalization of the adjacency matrix. The SGC model is defined as

$$
 X' = (\tilde{D}^{-\frac{1}{2}}\tilde{A}\tilde{D}^{-\frac{1}{2}})^K XW
$$

where $X$ is the node feature matrix, $W$ is the weight matrix, $\tilde{A}$ is the adjacency matrix with self-connections added, $\tilde{D}$ is the degree matrix of $\tilde{A}$, and $K$ is the number of layers.

\subsubsection{Graph Attention Networks (GAT)}

Graph Attention Networks (GAT)~\cite{velickovic2020pointer} are a class of neural networks that use attention mechanisms to aggregate information from neighboring nodes. The GAT model is defined as

$$
    h_i^{(l+1)} = \sigma(\sum_{j\in N(i)}\alpha_{ij}^{(l)}W^{(l)}h_j^{(l)})
$$

where $h_i^{(l)}$ is the node feature vector of node $i$ at layer $l$, $W^{(l)}$ is the weight matrix of the layer, $\sigma$ is the activation function, and $\alpha_{ij}^{(l)}$ is the attention coefficient between node $i$ and node $j$ at layer $l$. The attention coefficient is computed as

\subsubsection{Message Passing Neural Networks (MPNN)}
Message Passing Neural Networks (MPNNs)~\cite{gilmer2017neural} has been the leading model for graph learning. The MPNN model is defined as

$$
    m_i^{(t+1)} = \sum_{j\in N(i)}M^{(t)}(h_i^{(t)}, h_j^{(t)}, e_{ij})
$$

where $m_i^{(t)}$ is the message vector of node $i$ at iteration $t$, $M^{(t)}$ is the message function, $h_i^{(t)}$ is the node feature vector of node $i$ at iteration $t$, $e_{ij}$ is the edge feature vector between node $i$ and node $j$, and $N(i)$ is the neighborhood of node $i$. The message function $M^{(t)}$ is a neural network that takes as input the node feature vectors of node $i$ and node $j$, and the edge feature vector between them, and outputs the message vector $m_i^{(t)}$.

\subsection{Graph parameters} \label{sec:graph_parameters}
Now we can introduce the concept of a graph parameter. A graph parameter is a function $f: \mathbb{G} \rightarrow \mathbb{R}^d$ for $d\in \mathbb{N} >0$, which maps a graph to a d-dimensional vector over the reals. Furthermore, such functions need to be permutation invariant, i.e., $f(G) = f(\pi(G))$ for any permutation $\pi$ of the nodes of $G$, where a permutation is a function $\pi: V \rightarrow V$ and $\pi(G)$ denotes the graph obtained by permuting the nodes of $G$ according to $\pi$. A graph parameter allows us to quantify certain properties of a graph, from the basic structural numbers (number of nodes, number of edges, average degree, etc.) to distance properties (average shortest path length, diameter, etc.) to more complex properties (graph clustering coefficient, average number of coloring in the 1-WL algorithm, etc.). This thesis will focus on seven graph parameters, with definitions provided in section~\ref{sec:graph_parameters}.

\section{Experimental study}

\subsection{Experimental setup}
The following section outlines the planned methods and materials to be employed in the experiment, including the training framework, the datasets, the models, and the graph parameters.

\subsubsection{Training Framework}
The training framework is based on PyTorch Geometric~\cite{fey2019fast}. PyTorch Geometric is a library for deep learning on graphs and other irregular structures. It consists of various methods and utilities to ease the implementation of GNNs.

\subsubsection{Hyperparameters tuning}
Before training the models, for each pair of dataset and model, the hyperparameters of the model are tuned. In the experiment, 5-fold cross-validation is used to tune the hyperparameters instead of 10-fold cross validation in the final experiment. The reason is that the 10-fold cross-validation is computationally expensive and time-consuming.

The hyperparameters include the learning rate, batch size, hidden dimension, number of epochs, patience in early stopping, number of hidden layers, type of normalization, and patience of plateau scheduler. Additionally, for the GATv2 model, the number of heads, the dropout rate, and residual are also tuned.

The hyperparameters are tuned using the Bayesian optimization method\cite{frazier2018tutorial}. The hyperparameters that result in the highest validation accuracy are selected as the optimal hyperparameters for the model.

There are five models and 57 datasets, resulting in 285 pairs of dataset and model. For each pair, the hyperparameters are tuned using the Bayesian optimization method. The hyperparameters are tuned using the validation dataset. The validation dataset is a subset of the training dataset that is used to tune the hyperparameters of the model. The hyperparameters that result in the highest validation accuracy are selected as the optimal hyperparameters for the model.

\subsubsection{Datasets}
The dataset used for the classification is TUDataset~\cite{morris2020tudataset}. TUDataset is often used for the GNN evaluation. It consists of data from different domains, including small molecules, bioinformatics, social networks and synthetic. Since the size of some datasets is quite small (less than 500 data points/graphs), I use 10-fold cross validation in the training process, in order to fully utilize the data. A dataset is split into 1:1:8, where one of the folds is treated as the test dataset and another is treated as the validation dataset. The remaining folds are used for training. The training is repeated 10 times, each time with a different test fold. The average of the generalization error over the 10 runs is calculated as the final generalization error of the model.

I list information about the 57 datasets used in the experiment in the appendix~\ref{sec:datasets}.

\subsubsection{Models}
In the experiment, I plan to use different GNN layers in the model, including Graph Convolutional Networks(GCN)~\cite{kipf2016semi}, Graph Attention Networks(GAT)~\cite{velickovic2020pointer}, Graph Attention Networks(GATv2)~\cite{brody2021attentive}, Simplified Graph Convolution(SGC)~\cite{wu2019simplifying}, and Message Passing Neural Networks(MPNN)~\cite{gilmer2017neural}. The GCN model is shown in the appendix. The GCN model consists of four GCN layers, each followed by a ReLU activation function. The output of the last GCN layer is passed through a global mean pooling layer, followed by two fully connected layers.

\subsubsection{Experimental details}
The Adam optimizer~\cite{kingma2014adam} is considered to be employed in the model. To prevent overfitting and to optimize the use of time, the early stopping is employed. The loss function is defined as the cross-entropy loss function. The hyperparameters of the model are considered to be tuned. The hyperparameters include the learning rate, batch size, hidden dimension, number of epochs, and patience in early stopping. The generalization error is calculated as the difference between the training accuracy and the test accuracy.

\subsubsection{Graph Parameters Under Investigation}
As graph parameters defined abstractly in the section~\ref{sec:graph_parameters}, I give here the concrete definition of the graph parameters that are used in the experiment. The graph parameters are calculated for each dataset and used as input features for the model. The graph parameters are calculated using the NetworkX library~\cite{hagberg2008exploring}.

\begin{description}
    \item [Average degree] The average degree of the graph $d_{avg}$ is defined as the average of the degrees of all nodes in the graph, that is
    $$ d_{avg} = \frac{1}{|V|} \sum_{v \in V} d(v)$$
    \item [Average shortest path length] The average shortest path length of the graph $a$ is defined as the average of the shortest paths between all pairs of nodes in the graph. The shortest path $d(s, t)$ between two nodes $v$ and $w$ is the minimum number of edges that need to be traversed to go from $v$ to $w$. The formal definition is
    $$
a = \frac{1}{|V|(|V|-1)} \sum_{v \in V} \sum_{w \in V, w \neq v} d(v, w)
    $$

    \item [Graph diameter] The graph diameter $d$ is the maximum of the shortest paths between all pairs of nodes in the graph, that is,
    $$
d = \max_{v \in V} \max_{w \in V, w \neq v} d(v, w)
    $$

    \item [Graph density] The graph density $p$ is the ratio of the number of edges in the graph to the number of the maximum possible edges in the graph, that is,
    $$
p = \frac{2|E|}{|V|(|V|-1)}
    $$

    \item [Graph clustering coefficient] The graph clustering coefficient $C$ is a measure of the degree to which nodes in a graph tend to cluster together. The clustering coefficient of a node $v$ (as known as the local clustering coefficient) is defined as the fraction of the number of triangles that include node $v$ to the maximum possible number of triples centered on node $v$. The clustering coefficient of the graph (as known as the global clustering coefficient) is defined as the average of the clustering coefficients of all nodes in the graph. The formal definition is
    $$
C = \frac{1}{|V|} \sum_{v \in V} \frac{2\cdot |\{(i,j)\in E \mid i, j \in N(v)\}|}{d(v)(d(v)-1)}
    $$
    where the term $|\{(i,j)\in E \mid i, j \in N(v)\}|$ is the number of edges between the neighbors of node $v$, i.e. the number of triangles containing node $v$. Furthermore, the term $\sum_{v \in V} \frac{2\cdot |\{(i,j)\in E \mid i, j \in N(v)\}|}{d(v)(d(v)-1)}$ is the sum of the local clustering coefficients of all nodes in the graph, in which the degree of node $v$ is denoted as $d(v)$.

    \item [Centrality measure of graphs] At first we have the definition of centrality measure in the \textbf{node level}. The centrality measure of a node is a measure of the importance of the node in the graph. There are many centrality measures, such as degree centrality, closeness centrality, betweenness centrality, and eigenvector centrality.

    \textit{Degree centrality} is identical to the average degree of the graph defined above.

    \textit{Closeness centrality} $C_C(v)$ is defined as the reciprocal of the sum of the length of the shortest paths from the node $v$ to all other nodes in the graph, that is 
    $$C_C(v) = \frac{1}{\sum_{w\in V}{d(v ,w)}}$$
    where $d(v, w)$ is the shortest path between node $v$ and node $w$.

    \textit{Betweenness centrality} $C_B(v)$ is the sum of the fraction of the shortest paths between all pairs of nodes that pass through node $v$, that is

    $$
        C_B(v) = \sum_{s,t \in V} \frac{\sigma_{st}(v)}{\sigma_{st}}
    $$
    where $\sigma_{st}$ is the number of the shortest paths between nodes $s$ and $t$, and $\sigma_{st}(v)$ is the number of the shortest paths between nodes $s$ and $t$ that pass through node $v$.

    \textit{Eigenvector centrality} is a measure of the influence of a node in a graph based on the concept that connections to high-scoring nodes contribute more to the score of the node in question. Formally, the eigenvector centrality $C_E(v)$ of a node $v$  is defined as the $v$-th component of the eigenvector corresponding to the largest eigenvalue of the adjacency matrix $A$ of the graph. Mathematically, it can be expressed as:

    \[ C_E(v) = \frac{1}{\lambda} \sum_{u \in N(v)} a_{vu} C_E(u) \]

    where $ \lambda$ is the largest eigenvalue of the adjacency matrix $A$.
$a_{vu}$ is the element of the adjacency matrix $A$ corresponding to the edge between nodes $v$  and $u$.
$ C_E(u)$  is the eigenvector centrality of node $u$.

    Finally, in the \textbf{graph level}, the centrality measure of the graph is defined as the average of the centrality measures over all nodes in the graph.

    \item [Average number of coloring in the 1-WL algorithm] 
    \sloppy
    The 1-dimensional Weisfeiler-Leman algorithm (1-WL)~\cite{weisfeiler1968reduction} is an algorithm that assigns a unique label(color) to each node in the graph, also known as color refinement. The average number of coloring in the 1-WL algorithm is defined as the average number of colors used to color the nodes in the graph. The 1-WL algorithm is a powerful graph isomorphism algorithm that can be used to determine if two graphs are isomorphic. I give the procedure of the 1-WL algorithm here:
    \begin{enumerate}
        \item Assign the same color to all nodes in the graph.
        \item Two nodes $v$, $u$ are assigned a different color if there is a color c such that the number of c-colored neighbors of $u$ is not equal to the number of c-colored neighbors of $v$.
        \item Repeat step 2 until the colors of all nodes do not change.
    \end{enumerate}

\end{description}

\subsection{Results}

\subsubsection{Graph parameters of the datasets}

\begin{table}[!ht]
    \footnotesize
    \centering
    \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
    \hline
        \textbf{Name} & \textbf{Ave. degree} & \textbf{Ave. shortest path} & \textbf{Graph diameter} & \textbf{Graph density} & \textbf{Graph clustering coefficient} & \textbf{Ave. closeness centrality} & \textbf{Ave. betweenness centrality} & \textbf{Ave. eigenvector centrality} & \textbf{1-WL color count} \\ \hline
        COIL-DEL & 4.673071633424515 & 2.128750184714701 & 4.06974358974359 & 0.3275202823564923 & 0.546258039207784 & 0.5122126871677409 & 0.0647750716657882 & 0.2431245565997526 & 21.102564102564106 \\ \hline
        COIL-RAG & 1.8277218368573067 & 1.0604823111763986 & 1.242159383033419 & 0.924030303030303 & 0.5903545713545713 & 0.94372526076617 & 0.0462785547785547 & 0.5905014648511226 & 1.3115384615384615 \\ \hline
        Cuneiform & 4.079588014981273 & 2.244427103288844 & 3.0 & 0.2222330604444725 & 0.8634154547075895 & 0.4613474974345972 & 0.0722909733763576 & 0.1963017855703223 & 2.0 \\ \hline
        Fingerprint & 1.4868577990824814 & 2.115136284036567 & 4.344413012729844 & 0.4391785351497472 & 0.0017307791737759 & 0.5069423734352098 & 0.1840341453794884 & 0.4207756269406634 & 3.656584457887389 \\ \hline
        FIRSTMM\_DB & 4.50236815941043 & 12.309044701937715 & 30.676056338028168 & 0.0074027477877271 & 0.2631922211539924 & 0.0551742654129266 & 0.0188189276256195 & 0.0092704028700914 & 1357.8292682926829 \\ \hline
        Letter-high & 1.889683607525296 & 1.4794456563750826 & 2.444490301279405 & 0.5791851851851852 & 0.2976163139329805 & 0.6794971082752952 & 0.1657867724867724 & 0.4611670600331147 & 2.994222222222222 \\ \hline
        Letter-low & 1.3213534599145254 & 1.33364428641149 & 1.955783308931186 & 0.4165470899470899 & 0.0 & 0.4955472438922087 & 0.1618904761904762 & 0.4407481624189365 & 2.471111111111111 \\ \hline
        Letter-med & 1.352705840084288 & 1.3502031822344325 & 2.006610576923077 & 0.4245777777777778 & 0.0139393298059964 & 0.5071406868429822 & 0.1683222222222222 & 0.4416562761104123 & 2.517333333333333 \\ \hline
        MSRC\_9 & 4.81534002269555 & 3.113629895894764 & 7.004524886877828 & 0.1235873024311543 & 0.5412981657910688 & 0.33102605797573 & 0.055215313789426 & 0.1368862401087341 & 40.36199095022624 \\ \hline
        MSRC\_21 & 5.102391318360188 & 4.094474312024695 & 9.621669626998225 & 0.0680571076593652 & 0.5146660459581984 & 0.251511488615992 & 0.041532868642658 & 0.0936632954729105 & 77.22735346358792 \\ \hline
        MSRC\_21C & 4.782574662751558 & 3.113761314016 & 6.971291866028708 & 0.1240107738760322 & 0.5427170070575877 & 0.3315153420789583 & 0.0557392711060186 & 0.1371850938080253 & 40.05263157894737 \\ \hline
        DD & 4.979061666490266 & 7.919898418826995 & 19.70016611295681 & 0.0278322962071925 & 0.4793770769559906 & 0.1402074694227141 & 0.0324963430907113 & 0.0342447661028087 & 278.28013582342953 \\ \hline
        ENZYMES & 3.862625308732192 & 4.2810541955964485 & 10.4265625 & 0.1598878125179543 & 0.4533912862690096 & 0.263573448964092 & 0.1158938852730127 & 0.1471447702872793 & 28.575 \\ \hline
        KKI & 3.1856769969664422 & 3.500870554422239 & 8.144578313253012 & 0.1792960793494088 & 0.3526436673677589 & 0.3358793648012939 & 0.1297799757896069 & 0.1757034314039908 & 24.81927710843373 \\ \hline
        OHSU & 4.327667933476122 & 4.921680537597515 & 12.69620253164557 & 0.0822944467200859 & 0.3808295357080805 & 0.2407028072344937 & 0.0655860979728088 & 0.0823969445758651 & 77.51898734177215 \\ \hline
        Peking\_1 & 3.5514470044304343 & 3.968106920104182 & 9.623529411764704 & 0.1276151076564966 & 0.3673053988324483 & 0.2889094228756093 & 0.0998385744518594 & 0.1284748145692827 & 36.72941176470588 \\ \hline
        PROTEINS & 3.734642166440163 & 4.602185787956082 & 11.297071129707112 & 0.2121756091042708 & 0.5141966922501331 & 0.3022857727840693 & 0.118375456251276 & 0.1695057643632408 & 34.46900269541779 \\ \hline
        PROTEINS\_full & 3.734642166440163 & 4.602185787956082 & 11.297071129707112 & 0.2121756091042708 & 0.5141966922501331 & 0.3022857727840693 & 0.118375456251276 & 0.1695057643632408 & 34.46900269541779 \\ \hline
        COLORS-3 & 2.928853359699249 & 2.5810313856214724 & 5.275766205587198 & 0.1695881211511339 & 0.1450387037596311 & 0.3502761440955542 & 0.0639043708084029 & 0.1737080119571242 & 56.95904761904762 \\ \hline
        SYNTHETIC & 3.920000076293945 & 3.5064646464646465 & 7.0 & 0.0395959595959595 & 0.0203095238095238 & 0.2893391429701345 & 0.0255761698618841 & 0.0831547830080052 & 100.0 \\ \hline
        SYNTHETICnew & 3.9250000715255737 & 3.4872567329242026 & 7.274834437086093 & 0.0396464646464646 & 0.0229497089947089 & 0.289441726009901 & 0.0255427746856318 & 0.0836273264155355 & 99.96333333333334 \\ \hline
        Synthie & 3.6210444450378416 & 2.6348743925562266 & 5.494845360824742 & 0.038420429009193 & 0.0937036201778145 & 0.2766177855177352 & 0.0231922369186012 & 0.0786632559723412 & 90.6125 \\ \hline
        TRIANGLES & 3.1404920884927114 & 2.263375484124684 & 4.557498816896771 & 0.2416135135324899 & 0.2286352011335792 & 0.4374025890339534 & 0.0821216496736849 & 0.226180376713217 & 19.88371111111111 \\ \hline
        AIDS & 2.0128653862178325 & 3.490507728804514 & 7.762604013705335 & 0.1935408261794486 & 0.0071746166779525 & 0.3500774185559109 & 0.1911961844012523 & 0.262770387548969 & 11.8145 \\ \hline
        BZR & 2.1466610419897387 & 5.154837461860052 & 11.65679012345679 & 0.0648920405178712 & 8.903397114779228e-05 & 0.2054846442932875 & 0.1245291263310628 & 0.1416530437212679 & 29.607407407407408 \\ \hline
        BZR\_MD & 20.30392156862745 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.0 & 0.2203726689661978 & 1.0 \\ \hline
        COX2 & 2.107735009469108 & 5.936420803578038 & 13.790149892933618 & 0.0529052472359257 & 0.0007216110850642 & 0.1757209022905502 & 0.1262987138392391 & 0.1288265342157682 & 28.09421841541756 \\ \hline
        COX2\_MD & 25.27722772277228 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.0 & 0.1957034193657646 & 1.0 \\ \hline
        DHFR & 2.102359890622437 & 6.144674814243727 & 14.603174603174605 & 0.0536754666970741 & 0.0 & 0.1734264353163348 & 0.1288500765571343 & 0.1271607693927877 & 32.21560846560847 \\ \hline
        DHFR\_MD & 22.8676844783715 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.0 & 0.2076620678810238 & 1.0 \\ \hline
        ER\_MD & 20.327354260089685 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.0 & 0.2244695039966339 & 1.0 \\ \hline
        FRANKENSTEIN & 2.0627741588132293 & 3.6742676066855062 & 8.423452029938762 & 0.1705038719060743 & 0.0106826329262756 & 0.3148146889108825 & 0.1923606317376928 & 0.2432099454364741 & 13.42448697256168 \\ \hline
        MCF-7 & 2.1516180606917765 & 5.050070569624072 & 12.141178938535766 & 0.0980572307942984 & 0.0020345164969653 & 0.2175444791564102 & 0.1724077067751488 & 0.1725022149133201 & 21.418113071660063 \\ \hline
        MCF-7H & 2.0888838372172147 & 5.858613378052236 & 13.306228605946709 & 0.0537227676449342 & 0.0005825022974507 & 0.1778419519669024 & 0.1198217621207925 & 0.1223362659022958 & 32.9756211739287 \\ \hline
        MOLT-4 & 2.146077948348501 & 4.951738679286625 & 11.874633536790505 & 0.0987082247303717 & 0.0019246147011501 & 0.2183304220273518 & 0.1714407445609207 & 0.1734293274657891 & 21.028995347667543 \\ \hline
        MOLT-4H & 2.0858484049375803 & 5.738056140581633 & 12.992424764134702 & 0.0539489355320176 & 0.0005343843281943 & 0.17815930623996 & 0.1192636780853368 & 0.1228930198480923 & 32.272249465610464 \\ \hline
        Mutagenicity & 2.0379886209469995 & 4.418912992995495 & 9.63701923076923 & 0.0912875415379717 & 0.0031125822886973 & 0.2466634939984061 & 0.1371754284356414 & 0.1710356482094367 & 20.874567673507038 \\ \hline
        MUTAG & 2.18877208359698 & 3.626150957182829 & 8.21808510638298 & 0.1384540671648269 & 0.0 & 0.293764054092514 & 0.1694459751944365 & 0.2197353385133972 & 14.8031914893617 \\ \hline
        NCI1 & 2.15501380214726 & 5.234873158476093 & 12.659402381487306 & 0.088944910671084 & 0.0030999008340164 & 0.2026186665583615 & 0.1631676786426253 & 0.1607325923179734 & 24.721411192214116 \\ \hline
        NCI109 & 2.156446177386937 & 5.183950502801717 & 12.49528936742934 & 0.0894286690245151 & 0.0030732389355228 & 0.20411132911515 & 0.1624231134845694 & 0.1617585552708976 & 24.52410952265568 \\ \hline
        NCI-H23 & 2.1451602406810686 & 4.950946501654924 & 11.869525050337558 & 0.0987776628128759 & 0.0018996992922189 & 0.2183535143438838 & 0.1715792456717554 & 0.1735208031755432 & 20.99303645329963 \\ \hline
        NCI-H23H & 2.085327596320919 & 5.734343001278942 & 12.983736711151325 & 0.0539842770435475 & 0.0005268942359655 & 0.1782099752168056 & 0.1192891041739607 & 0.1229635264798962 & 32.22578246970485 \\ \hline
        OVCAR-8 & 2.145637268375766 & 4.945169149304127 & 11.85222800857742 & 0.0987730040211328 & 0.0018986224960467 & 0.2183941696064956 & 0.171403631603811 & 0.1735393405097124 & 20.98556125974924 \\ \hline
        OVCAR-8H & 2.0856259705836995 & 5.728655397369023 & 12.966769410758882 & 0.0539777071171825 & 0.0005298176630677 & 0.1782129665549435 & 0.1192148511162624 & 0.1229665170804826 & 32.22561457202093 \\ \hline
        P388 & 2.109217112454913 & 4.355752638151867 & 10.21824835544779 & 0.1190787859581164 & 0.0037544697177687 & 0.2459957278657192 & 0.1789538865626834 & 0.1958220277248745 & 17.133174189814813 \\ \hline
        P388H & 2.0638945394922854 & 5.289795718220624 & 11.765664188214403 & 0.0639638393366292 & 0.0009642194122106 & 0.1971310478874152 & 0.1240122910178544 & 0.1373237905541068 & 26.71100983796296 \\ \hline
        PC-3 & 2.1520653447146643 & 5.05018085139685 & 12.140862989009436 & 0.098055200309084 & 0.0019733135754618 & 0.2175815805232795 & 0.1725492873407526 & 0.172560720893664 & 21.37340506743248 \\ \hline
        PC-3H & 2.08922384690422 & 5.857822430952191 & 13.30650249313117 & 0.0537656750613318 & 0.0005561305149752 & 0.1779189564869681 & 0.1200247359820712 & 0.1224665687629844 & 32.881929550328984 \\ \hline
        PTC\_FM & 1.9758043374578045 & 3.311647990381311 & 7.36676217765043 & 0.2154944740700171 & 0.0091220986824968 & 0.3693003427913903 & 0.2055882581121011 & 0.2786143302706905 & 9.91404011461318 \\ \hline
        PTC\_FR & 1.9863111707899304 & 3.387662268892836 & 7.589743589743589 & 0.2085568638656639 & 0.0088214579442649 & 0.3612672957602951 & 0.2045358868174996 & 0.2734008790739162 & 10.22222222222222 \\ \hline
        PTC\_MM & 1.9701179610121815 & 3.301432053360394 & 7.339285714285714 & 0.2223452925857377 & 0.0130582137161084 & 0.374488207420213 & 0.2059637189235834 & 0.2823441282735879 & 9.803571428571429 \\ \hline
        PTC\_MR & 1.9805375618297 & 3.3577641223936463 & 7.523255813953488 & 0.2137888896192228 & 0.009509685286154 & 0.3666473296924281 & 0.2062170509277356 & 0.2777684525909222 & 10.040697674418604 \\ \hline
        SF-295 & 2.145489384462884 & 4.9445677038396125 & 11.852116182572614 & 0.09881526807439 & 0.0019131887956769 & 0.2184346701016761 & 0.1714761059761314 & 0.1735919743359789 & 20.97159246107621 \\ \hline
        SF-295H & 2.0855234748235936 & 5.727434951743517 & 12.9664274556856 & 0.0539895579758234 & 0.0005299113307649 & 0.1782697110808576 & 0.1192036786794211 & 0.1229816733300816 & 32.196146110104046 \\ \hline
        SN12C & 2.145026913904858 & 4.94825141099449 & 11.86401279205747 & 0.0987936439503507 & 0.001927945706436 & 0.2183468315283276 & 0.1714975830848956 & 0.1735570013676089 & 20.99167583241676 \\ \hline
        SN12CH & 2.0853053252079072 & 5.730086714838555 & 12.973598487437071 & 0.0540104189866913 & 0.0005390487212868 & 0.1781938757210136 & 0.1192686561921279 & 0.1230077396296653 & 32.236301369863014 \\ \hline
        SW-620 & 2.1452310279340394 & 4.94393636762053 & 11.851577038136291 & 0.0988746005232462 & 0.0018522578567537 & 0.2185003794708548 & 0.1714920826604129 & 0.1736489333480562 & 20.97197276226192 \\ \hline
        SW-620H & 2.0854275176564645 & 5.725806234940699 & 12.962851360467733 & 0.054054544610838 & 0.0005110193017508 & 0.1783225632271403 & 0.1192839052512742 & 0.1230820111720522 & 32.19651139840126 \\ \hline
        UACC257 & 2.145698715720169 & 4.955834964420495 & 11.885704040162564 & 0.0987250195703296 & 0.0018738116805416 & 0.218300263582104 & 0.1716221375949755 & 0.1734631006815176 & 20.994198259477844 \\ \hline
        UACC257H & 2.0856917230912635 & 5.740515645090362 & 13.003174313183676 & 0.0539749990285044 & 0.0005209407683445 & 0.1781806662974605 & 0.1193640515072574 & 0.122956068725157 & 32.22436731019306 \\ \hline
        Yeast & 2.099442849357738 & 4.364322361013096 & 10.26343257027636 & 0.1210553127513687 & 0.0022671334945899 & 0.2476806589296417 & 0.183060827627568 & 0.1985665767951158 & 16.67475282973832 \\ \hline
        YeastH & 2.0589509621359707 & 5.245391698622407 & 11.6875493153864 & 0.0642704732457107 & 0.0005773570062507 & 0.1983872537891083 & 0.1254655281986756 & 0.1384817929367545 & 26.05320284921044 \\ \hline
        Sum & 242.8347529362711 & 258.6723790469145 & 587.9183294923151 & 12.731666924394474 & 11.845921907393928 & 20.687237238791475 & 7.3848149134808 & 11.636469235194477 & 3245.599253687252 \\ \hline
    \end{tabular}
    \caption{Graph parameters of the datasets}
\end{table}

\newpage

\subsubsection{Generalization error of GCN model}

\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
        Name & Ave. generalization error & Standard deviation \\ \hline
        MCF-7 & 0.002249991986900568 & 0.01604018546640873 \\ \hline
        MCF-7H & 0.004124999046325684 & 0.02702830731868744 \\ \hline
        MOLT-4 & 0.0 & 0.014102176763117313 \\ \hline
        MOLT-4H & 0.008624988608062267 & 0.01870434731245041 \\ \hline
        Mutagenicity & 0.03765624016523361 & 0.024758048355579376 \\ \hline
        MUTAG & 0.13552632927894592 & 0.10957122594118118 \\ \hline
        NCI1 & 0.06834375858306885 & 0.0339551605284214 \\ \hline
        NCI109 & 0.07212500274181366 & 0.030083369463682175 \\ \hline
        NCI-H23 & 0.003218746278434992 & 0.013211547397077084 \\ \hline
        NCI-H23H & 0.0006250202422961593 & 0.011396590620279312 \\ \hline
        OVCAR-8 & -0.0001249849738087505 & 0.015103230252861977 \\ \hline
        OVCAR-8H & 0.0001562476099934429 & 0.013315632939338684 \\ \hline
        P388 & 0.008062511682510376 & 0.01043933816254139 \\ \hline
        P388H & 0.006343752145767212 & 0.007148966193199158 \\ \hline
        PC-3 & 0.0008750140550546348 & 0.012619029730558395 \\ \hline
        PC-3H & 0.0001250088243978098 & 0.00780930370092392 \\ \hline
        PTC\_FM & 0.12454469501972198 & 0.14246851205825806 \\ \hline
        PTC\_FR & 0.08159632235765457 & 0.09593729674816132 \\ \hline
        PTC\_MM & 0.12555554509162903 & 0.12687508761882782 \\ \hline
        PTC\_MR & 0.058610398322343826 & 0.11794959753751755 \\ \hline
        SF-295 & 0.0009687542915344238 & 0.013665653765201569 \\ \hline
        SF-295H & 0.004500013776123524 & 0.018203353509306908 \\ \hline
        SN12C & 0.0004999935626983643 & 0.017035124823451042 \\ \hline
        SN12CH & 0.006718760821968317 & 0.016944456845521927 \\ \hline
        SW-620 & 5.9604645663569045e-09 & 0.006674798205494881 \\ \hline
        SW-620H & 0.002499997615814209 & 0.008080806583166122 \\ \hline
        UACC257 & 0.0022812604438513517 & 0.013034629635512829 \\ \hline
        UACC257H & 6.25073880655691e-05 & 0.015311303548514843 \\ \hline
        Yeast & 0.0022187530994415283 & 0.01859860308468342 \\ \hline
        YeastH & 0.00024999381275847554 & 0.015245940536260605 \\ \hline
        PROTEINS\_full & 0.11115662008523941 & 0.5160216689109802 \\ \hline
        DD & -0.1750326007604599 & 0.3490835428237915 \\ \hline
        ENZYMES & 0.3110416531562805 & 0.06384395807981491 \\ \hline
        KKI & 0.04776120185852051 & 0.2546080946922302 \\ \hline
        OHSU & 0.39824172854423523 & 0.138640895485878 \\ \hline
        Peking\_1 & 0.2980072498321533 & 0.2580110728740692 \\ \hline
        COLORS-3 & 0.07568749040365219 & 0.03585537523031235 \\ \hline
        SYNTHETIC & -0.41749995946884155 & 0.3929196000099182 \\ \hline
        SYNTHETICnew & -0.36250001192092896 & 0.45766785740852356 \\ \hline
        Synthie & 0.14562499523162842 & 0.032582689076662064 \\ \hline
    \end{tabular}
    \caption{Generalization error of the GCN model on 57 datasets}
    \label{tab:ge_gcn} % ge -> generalization error
\end{table}

\newpage

\subsubsection{Generalization error of SGC model}
\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
        Name & Ave. generalization error & Standard deviation \\ \hline
        PROTEINS\_full & 0.04741409420967102 & 0.5624663829803467 \\ \hline
        DD & -0.21720178425312042 & 0.40132829546928406 \\ \hline
        ENZYMES & 0.3279166519641876 & 0.060605552047491074 \\ \hline
        KKI & 0.10951493680477142 & 0.24877962470054626 \\ \hline
        OHSU & 0.07604396343231201 & 0.1943538784980774 \\ \hline
        Peking\_1 & 0.1769927442073822 & 0.22724425792694092 \\ \hline
        COLORS-3 & 0.11781251430511475 & 0.04669201001524925 \\ \hline
        SYNTHETIC & -0.37541666626930237 & 0.4494529068470001 \\ \hline
        SYNTHETICnew & -0.36666664481163025 & 0.46157106757164 \\ \hline
        Synthie & 0.13749998807907104 & 0.042466290295124054 \\ \hline
        AIDS & 0.006437480449676514 & 0.011520843021571636 \\ \hline
        BZR & 0.10815385729074478 & 0.21917256712913513 \\ \hline
        BZR\_MD & 0.1678861677646637 & 0.09830881655216217 \\ \hline
        COX2 & 0.07495072484016418 & 0.13212721049785614 \\ \hline
        COX2\_MD & 0.14868313074111938 & 0.16540752351284027 \\ \hline
        DHFR & 0.1332409381866455 & 0.1551705300807953 \\ \hline
        DHFR\_MD & 0.0777045264840126 & 0.2503173351287842 \\ \hline
        ER\_MD & -0.008367183618247509 & 0.15631815791130066 \\ \hline
        FRANKENSTEIN & 0.10787498950958252 & 0.05383045971393585 \\ \hline
        MCF-7 & 0.35490626096725464 & 0.17573890089988708 \\ \hline
        MCF-7H & 0.3593124747276306 & 0.1884346306324005 \\ \hline
        MOLT-4 & 0.18021871149539948 & 0.39803561568260193 \\ \hline
        MOLT-4H & 0.18649999797344208 & 0.3838197886943817 \\ \hline
        Mutagenicity & 0.04325000196695328 & 0.028246337547898293 \\ \hline
        MUTAG & 0.07375730574131012 & 0.13988283276557922 \\ \hline
        NCI1 & 0.09100000560283661 & 0.16161812841892242 \\ \hline
        NCI109 & -0.06128125637769699 & 0.2943810522556305 \\ \hline
        NCI-H23 & 0.31696873903274536 & 0.07884199917316437 \\ \hline
        NCI-H23H & 0.3696562349796295 & 0.06490368396043777 \\ \hline
        OVCAR-8 & 0.3291874825954437 & 0.0815950483083725 \\ \hline
        OVCAR-8H & 0.2902187705039978 & 0.0771569237112999 \\ \hline
        P388 & 0.3618437647819519 & 0.13428252935409546 \\ \hline
        P388H & -0.052218757569789886 & 0.3646376132965088 \\ \hline
        PC-3 & -0.08028124272823334 & 0.29225972294807434 \\ \hline
        PC-3H & 0.2593437135219574 & 0.22314861416816711 \\ \hline
        PTC\_FM & 0.10863511264324188 & 0.14730817079544067 \\ \hline
        PTC\_FR & 0.08632432669401169 & 0.09645862877368927 \\ \hline
        PTC\_MM & 0.09649830311536789 & 0.0817335695028305 \\ \hline
        PTC\_MR & 0.060720354318618774 & 0.18291200697422028 \\ \hline
        SF-295 & 0.3124375343322754 & 0.11770480871200562 \\ \hline
        SF-295H & 0.28881245851516724 & 0.12331245839595795 \\ \hline
        SN12C & 0.2917812764644623 & 0.05537300556898117 \\ \hline
        SN12CH & 0.3397812247276306 & 0.09572991728782654 \\ \hline
        SW-620 & 0.3427812457084656 & 0.2430398166179657 \\ \hline
        SW-620H & 0.39274999499320984 & 0.1998475342988968 \\ \hline
        UACC257 & -0.10218751430511475 & 0.2883416414260864 \\ \hline
        UACC257H & 0.3322499692440033 & 0.22775331139564514 \\ \hline
        Yeast & 0.0 & 0.0 \\ \hline
        YeastH & 0.0 & 0.0 \\ \hline
    \end{tabular}
\end{table}

\subsubsection{Generalization error of GATv2 model}

\subsubsection{Generalization error of MPNN model}

\subsubsection{Corrolation analysis between generalization error and graph parameters}

\subsection{Discussion}

\section{Conclusion}

\clearpage

\bibliographystyle{abbrvnat}
\bibliography{main} % refers to example.bib

\clearpage

\section{Appendix}

\subsection{Basic information about TUDataset}\label{sec:datasets}
This table~\ref{tab:datasets} shows the information about a list of datasets used in this thesis.

\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|l|l|l|l|}
    \hline
        name & Graphs & Classes & Avg. Nodes & Avg. Edges & catagory \\ \hline
        AIDS & 2000 & 2 & 15.69 & 16.20 & Small molecules \\ \hline
        BZR & 405 & 2 & 35.75 & 38.36 & Small molecules \\ \hline
        BZR\_MD & 306 & 2 & 21.30 & 225.06 & Small molecules \\ \hline
        COX2 & 467 & 2 & 41.22 & 43.45 & Small molecules \\ \hline
        COX2\_MD & 303 & 2 & 26.28 & 335.12 & Small molecules \\ \hline
        DHFR & 756 & 2 & 42.43 & 44.54 & Small molecules \\ \hline
        DHFR\_MD & 393 & 2 & 23.87 & 283.02 & Small molecules \\ \hline
        ER\_MD & 446 & 2 & 21.33 & 234.85 & Small molecules \\ \hline
        FRANKENSTEIN & 4337 & 2 & 16.90 & 17.88 & Small molecules \\ \hline
        MCF-7 & 27770 & 2 & 26.40 & 28.53 & Small molecules \\ \hline
        MCF-7H & 27770 & 2 & 47.31 & 49.44 & Small molecules \\ \hline
        MOLT-4 & 39765 & 2 & 26.10 & 28.14 & Small molecules \\ \hline
        MOLT-4H & 39765 & 2 & 46.70 & 48.74 & Small molecules \\ \hline
        Mutagenicity & 4337 & 2 & 30.32 & 30.77 & Small molecules \\ \hline
        MUTAG & 188 & 2 & 17.93 & 19.79 & Small molecules \\ \hline
        NCI1 & 4110 & 2 & 29.87 & 32.30 & Small molecules \\ \hline
        NCI109 & 4127 & 2 & 29.68 & 32.13 & Small molecules \\ \hline
        NCI-H23 & 40353 & 2 & 26.07 & 28.10 & Small molecules \\ \hline
        NCI-H23H & 40353 & 2 & 46.67 & 48.70 & Small molecules \\ \hline
        OVCAR-8 & 40516 & 2 & 26.08 & 28.11 & Small molecules \\ \hline
        OVCAR-8H & 40516 & 2 & 46.67 & 48.70 & Small molecules \\ \hline
        P388 & 41472 & 2 & 22.11 & 23.56 & Small molecules \\ \hline
        P388H & 41472 & 2 & 40.45 & 41.89 & Small molecules \\ \hline
        PC-3 & 27509 & 2 & 26.36 & 28.49 & Small molecules \\ \hline
        PC-3H & 27509 & 2 & 47.20 & 49.33 & Small molecules \\ \hline
        PTC\_FM & 349 & 2 & 14.11 & 14.48 & Small molecules \\ \hline
        PTC\_FR & 351 & 2 & 14.56 & 15.00 & Small molecules \\ \hline
        PTC\_MM & 336 & 2 & 13.97 & 14.32 & Small molecules \\ \hline
        PTC\_MR & 344 & 2 & 14.29 & 14.69 & Small molecules \\ \hline
        SF-295 & 40271 & 2 & 26.06 & 28.09 & Small molecules \\ \hline
        SF-295H & 40271 & 2 & 46.65 & 48.68 & Small molecules \\ \hline
        SN12C & 40004 & 2 & 26.08 & 28.11 & Small molecules \\ \hline
        SN12CH & 40004 & 2 & 46.69 & 48.72 & Small molecules \\ \hline
        SW-620 & 40532 & 2 & 26.06 & 28.09 & Small molecules \\ \hline
        SW-620H & 40532 & 2 & 46.63 & 48.66 & Small molecules \\ \hline
        UACC257 & 39988 & 2 & 26.09 & 28.13 & Small molecules \\ \hline
        UACC257H & 39988 & 2 & 46.68 & 48.71 & Small molecules \\ \hline
        Yeast & 79601 & 2 & 21.54 & 22.84 & Small molecules \\ \hline
        YeastH & 79601 & 2 & 39.45 & 40.75 & Small molecules \\ \hline
        DD & 1178 & 2 & 284.32 & 715.66 & Bioinformatics \\ \hline
        ENZYMES & 600 & 6 & 32.63 & 62.14 & Bioinformatics \\ \hline
        KKI & 83 & 2 & 26.96 & 48.42 & Bioinformatics \\ \hline
        OHSU & 79 & 2 & 82.01 & 199.66 & Bioinformatics \\ \hline
        Peking\_1 & 85 & 2 & 39.31 & 77.35 & Bioinformatics \\ \hline
        PROTEINS\_full & 1113 & 2 & 39.06 & 72.82 & Bioinformatics \\ \hline
        COIL-DEL & 3900 & 100 & 21.54 & 54.24 & Computer vision \\ \hline
        COIL-RAG & 3900 & 100 & 3.01 & 3.02 & Computer vision \\ \hline
        Cuneiform & 267 & 30 & 21.27 & 44.80 & Computer vision \\ \hline
        Fingerprint & 2149 & 15 & 7.06 & 5.76 & Computer vision \\ \hline
        FIRSTMM\_DB & 41 & 11 & 1377.27 & 3074.10 & Computer vision \\ \hline
        Letter-high & 2250 & 15 & 4.67 & 4.50 & Computer vision \\ \hline
        Letter-low & 2250 & 15 & 4.68 & 3.13 & Computer vision \\ \hline
        Letter-med & 2250 & 15 & 4.67 & 3.21 & Computer vision \\ \hline
        MSRC\_9 & 221 & 8 & 40.58 & 97.94 & Computer vision \\ \hline
        MSRC\_21 & 563 & 20 & 77.52 & 198.32 & Computer vision \\ \hline
        MSRC\_21C & 209 & 20 & 40.28 & 96.60 & Computer vision \\ \hline
        COLORS-3 & 10500 & 11 & 61.31 & 91.03 & Synthetic \\ \hline
        SYNTHETIC & 300 & 2 & 100.00 & 196.00 & Synthetic \\ \hline
        SYNTHETICnew & 300 & 2 & 100.00 & 196.25 & Synthetic \\ \hline
        Synthie & 400 & 4 & 95.00 & 172.93 & Synthetic \\ \hline
    \end{tabular}
    \caption{Basic information about the datasets used in the experiment}
    \label{tab:datasets}
\end{table}

\subsection{Example of GCN model}

\begin{lstlisting}[language=Python]
import torch

\end{lstlisting}

\end{document}