### 2025-01-29 paper/includes/abstract.tex

% !TeX spellcheck = en_US
% !TeX encoding = UTF-8
\chapter*{Abstract}
The goal of this bachelor thesis is to understand empirically how graph parameters/characteristics influence the generalization error of Graph Neural Networks (GNNs).