### 2025-02-20 paper/includes/abstract.tex

% !TeX spellcheck = en_US
% !TeX encoding = UTF-8
\chapter*{Abstract}

This thesis investigates the impact of various graph parameters on the generalization error of different Graph Neural Network (GNN) architectures. Extensive experiments were conducted on 49 graph classification datasets from TUDataset using four GNN models: Graph Convolutional Networks (GCN), Simplified Graph Convolution (SGC), Graph Attention Networks v2 (GATv2), and Message Passing Neural Networks (MPNN). The study examines graph parameters such as average degree, average clustering coefficient, average shortest path length, graph diameter, graph density, centrality measures, and 1-WL color count. The findings provide valuable insights into the relationship between graph parameters and the generalization performance of GNNs, informing the selection of appropriate GNN architectures for various graph parameters. Future work will focus on expanding the range of GNN architectures and datasets, as well as generating synthetic datasets with controlled properties to further investigate the impact of graph parameters on GNN generalization.